---
title: "p8105_hw3_jp3665"
author: "Joana Petrescu"
date: "12/10/2020"
output: github_document
---

```{r}

library(tidyverse)
library(p8105.datasets)
library(janitor)
library(knitr)
library(gridExtra)

```

## Instacart data

Load and explore data

```{r}

data("instacart") %>% tidy_names()

dim(instacart)

```

The "Instacart" data consist of 1384617 observations for 15 variables. The variables are as follows:

* `order_id`: order identifier
* `product_id`: product identifier
* `add_to_cart_order`: order in which each product was added to cart
* `reordered`: 1 if this product has been ordered by this user in the past, 0 otherwise
* `user_id`: customer identifier
* `order_number`: the order sequence number for this user (1=first, n=nth)
* `order_dow`: the day of the week on which the order was placed
* `order_hour_of_day`: the hour of the day on which the order was placed
* `days_since_prior_order`: days since the last order, capped at 30, NA if order_number=1
* `product_name`: name of the product
* `aisle_id:` aisle identifier
* `department_id`: department identifier
* `aisle`: the name of the aisle
* `department`: the name of the department

Here is a random observation from the data set as an example:

```{r}

instacart[61,]

```

This observation refers to organic garlic (produce department, fresh vegetables aisle) that was the 29th item to be added to cart on Tuesday at 8:00 as part of order #98 by user #56463. This user has ordered 40 times previously, they placed their previoous order 14 days prior to this one, and they have ordered this item before.

```{r}

instacart %>% count(aisle) %>% nrow()

instacart %>% count(aisle) %>% arrange(desc(n)) %>% head() %>% knitr::kable()

```

There are 134 unique aisles defined in this data set. This table demonstrates that the most items are ordered from the fresh vegetables and fresh fruits aisles.

Plot the number of items ordered from each isle for aisles with >10k items ordered.

```{r}

instacart %>% 
  count(aisle) %>%
  mutate(aisle = fct_reorder(aisle, n)) %>%
  filter(n > 10000) %>% 
  ggplot(aes(x = aisle, y = n)) +
  geom_bar(stat = "identity") +
  theme(
    axis.text.x = element_text(angle = 90)
  ) +
  xlab("Aisle") +
  ylab("Number of items") +
  ggtitle("Number of items ordered for each aisle")
  

```

Make a tablee showing the three most popular items in each of the following: baking ingredients, dog food care,  and packaged vegetables fruits.

```{r}

instacart %>%
  select(product_name, aisle) %>%
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) %>%
  group_by(aisle) %>%
  count(product_name) %>%
  top_n(3) %>%
  knitr::kable()

```


Make a table showing the mean hour of the day that Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week.

```{r}

instacart %>%
  select(product_name, order_hour_of_day, order_dow) %>%
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) %>%
  group_by(product_name, order_dow) %>%
  summarize(mean(order_hour_of_day)) %>%
  spread(order_dow, "mean(order_hour_of_day)") %>%
  rename(
    sunday = "0",
    monday = "1",
    tuesday = "2",
    wednesday = "3",
    thursday = "4",
    friday = "5",
    saturday = "6"
  ) %>%
  knitr::kable()

```

## Accelerometer data

This data set consists of accelerometer measurements collected over 5 weeks from a 63 year old male w/ BMI 25 who was admitted to the Advanced Cardiac Care Center at CUMC and diagnosed with CHF. The `activity.*` variables are activity counts for each minute of 24-hour day.

Load and tidy data. The final dataset includes

```{r}

accel_data <- read.csv("./data/accel_data.csv") %>%
  tidy_names() %>%
  mutate(
    weekend = as.logical(day %in% c("Saturday", "Sunday"))
  )

accel_data$weekend

activity_cols <- accel_data[,grepl("activity", names(accel_data))]

accel_data %>%
  mutate(
    total_activity = activity_cols %>% rowSums(na.rm = TRUE)
) %>%
  select(week, day_id, day, weekend, total_activity) %>%
  knitr::kable()



```

It seems that the days with the lowest total activity are weekends.

```{r}

accel_data %>% 
  gather("minute", "activity", 4:1443) %>%
  ggplot(aes(x = minute, y = activity)) +
           geom_line(aes(color = day)) +
  theme(axis.text.x = element_blank(),
        axis.ticks.x=element_blank())
  

```

From this plot, it actually appears that the most activity is in the morning and in the evening and that the subject had the most activity on Saturdays and Sundays and the least on Wednesdays.

## Weather data

Load and tidy data, make temperature values in degrees and precipitation values in mm. The original data has 2595176 observations of 7 variables including station ID, date, precipitation, snowfall, max temp, min temp. There are 3387623 NA values which suggests there is a lot of missing data.

```{r}

data("ny_noaa")

ny_noaa %>% is.na() %>% sum()

weather_data <- ny_noaa %>%
  mutate(year = lubridate::year(date), 
         month = lubridate::month(date), 
         day = lubridate::day(date),
         tmin = as.numeric(tmin) / 10,
         tmax = as.numeric(tmax) / 10,
         prcp = as.numeric(prcp) / 10)

weather_data %>% count(snow) %>% arrange(desc(n))

```

The most commonly observed values for snowfall are 0 and NA because it does not snow most days and there must be some inconsistencies with data entry (0 vs. NA).

Plot maximum temperatures for each station in January and July across years.

```{r}

p1 <- weather_data %>% 
  filter(month == "1") %>%
  select(id, tmax, date, year, month) %>%
  drop_na() %>%
  group_by(id) %>%
  ggplot(aes(x = date, y = tmax)) +
  geom_line(aes(color = id)) +
  theme(legend.position = "none") +
  ggtitle("Max temp in January (degrees)")

p2 <-  weather_data %>% 
  filter(month == "7") %>%
  select(id, tmax, date, year, month) %>%
  drop_na() %>%
  group_by(id) %>%
  ggplot(aes(x = date, y = tmax)) +
  geom_line(aes(color = id)) +
  theme(legend.position = "none") +
  ggtitle("Max temp in July (degrees)")

grid.arrange(p1, p2, ncol=2)

```
There is no obvious change over these years in maximum temperature although it does tend to fluctuate over time. Years 2005 and 1995 seem to be outliers of high temp in January and 1989, 1991, 2000, and 2006 seem to be outliers of low temps in July.

Make a plot showing maximum vs. minimum temperatures for the full dataset.

```{r}

p3 <- weather_data %>% 
  select(id, tmax, tmin, date, year, month) %>%
  ggplot(aes(x = date)) +
  geom_line(aes(x = date, y = tmax, color = "red")) +
  geom_line(aes(x = date, y = tmin, color = "blue")) +
  theme(legend.position = "none") +
  ggtitle("Max and min temp (degrees)")

p4 <- weather_data %>%
  select(year, snow) %>%
  filter(snow == (0:100)) %>%
  ggplot(aes(x = factor(year), y = snow)) +
  geom_violin() +
  theme(
    axis.text.x = element_text(angle = 90)
  ) +
  ggtitle("Snowfall by year (mm)")

grid.arrange(p3, p4, ncol=2)

```

