---
title: "p8105_hw3_jp3665"
author: "Joana Petrescu"
date: "12/10/2020"
output: github_document
---

```{r}

library(tidyverse)
library(p8105.datasets)
library(janitor)

```

## Instacart data

Load and explore data

```{r}

data("instacart") %>% tidy_names()

dim(instacart)

```

The "Instacart" data consist of 1384617 observations for 15 variables. The variables are as follows:

* `order_id`: order identifier
* `product_id`: product identifier
* `add_to_cart_order`: order in which each product was added to cart
* `reordered`: 1 if this product has been ordered by this user in the past, 0 otherwise
* `user_id`: customer identifier
* `order_number`: the order sequence number for this user (1=first, n=nth)
* `order_dow`: the day of the week on which the order was placed
* `order_hour_of_day`: the hour of the day on which the order was placed
* `days_since_prior_order`: days since the last order, capped at 30, NA if order_number=1
* `product_name`: name of the product
* `aisle_id:` aisle identifier
* `department_id`: department identifier
* `aisle`: the name of the aisle
* `department`: the name of the department

Here is a random observation from the data set as an example:

```{r}

instacart[61,]

```

This observation refers to organic garlic (produce department, fresh vegetables aisle) that was the 29th item to be added to cart on Tuesday at 8:00 as part of order #98 by user #56463. This user has ordered 40 times previously, they placed their previoous order 14 days prior to this one, and they have ordered this item before.

